---
title: "Feature_selection"
author: "Bundi Kirimi"
date: "2/5/2022"
output: html_document
---


#Feature Selection
http://bit.ly/CarreFourDataset






#Loading the data
```{r}


library(readr)
df <- read.csv("Supermarket_Dataset_1 - Sales Data.csv",sep = ",", dec = ".")

head(df)
```


#Previewing the summary of the dataset
```{r}
summary(df)
```

#Checking the data
```{r}
#Length
length(df)
```
The dataframe has 18 columns


````{r}
#Dimensions
dim(df)

````




#The dataframe has 12330 row entries and 18 columns
Column Names
```{r}
colnames(df)

```

Column data types

```{r}
sapply(df, class)
```




#Data Cleaning

```{r}

any(is.na(df))
```

NO NULL VALUES IN OUR DATA

#Data Types
```{r}

sapply(df, class)
```
Checking the categories in categorical columns
```{r}
unique(df$Branch)

unique(df$Customer.type)

unique(df$Gender)

unique(df$Product.line )
unique(df$Payment)

unique(df$ Rating)


```

```{r}
# Convert data types using as.integer
# Branch
# Convert data types using as.integer
# Branch
df$Branch_E<-as.integer(as.factor(df$Branch))
# Customer Type
df$Customer_Type_E<-as.integer(as.factor(df$Customer.type))
# Gender
df$Gender_E<-as.integer(as.factor(df$Gender))
# Product.line
df$Product_Line_E<-as.integer(as.factor(df$Product.line))
#Payment
df$Payment_E<-as.integer(as.factor(df$Payment))
```

Splitting Date_Time
```{r}
library("lubridate")
```



```{r}
library("ymd")

# Split date year, month and day.
# Convert to date datatype first then split thereafter
df$Date <- as.Date(df$Date, "%m/%d/%Y")
df$year <- year(ymd(df$Date))
df$month <- month(ymd(df$Date))
df$day <- day(ymd(df$Date))
df$hour = format(strptime(df$Time,"%H:%M"),'%H')
df$minute = format(strptime(df$Time,"%H:%M"),'%M')
```

```{r}
#install.packages(dplyr)
library(dplyr)
```

```{r}
df_sales_num <- select_if(df,is.numeric)
str(df_sales_num)
```






Identifying colums variance

A variance of zero indicates that all of the data values are identical. All non-zero variances are positive. A small variance indicates that the data points tend to be very close to the mean, and to each other.
We will drop the colum with zero variance as its not necessarry in our analysis because all the values are identical
```{r}

# Identify the columns with zero column variance.
names(df_sales_num[, sapply(df_sales_num, function(v) var(v,
na.rm=TRUE)==0)])


```
```{r}
# Drop the columns as they have a variance of 0"
df_sales_num <- subset(df_sales_num, select = -c(gross.margin.percentage,
year))
```


```{r}
dim(df_sales_num)
```










```{r}
# Installing and loading our caret package
# ---
# 
suppressWarnings(
        suppressMessages(if
                         (!require(caret, quietly=TRUE))
                install.packages("caret")))
library(caret)

```


```{R}
# Installing and loading the corrplot package for plotting
# ---
# 
suppressWarnings(
        suppressMessages(if
                         (!require(corrplot, quietly=TRUE))
                install.packages("corrplot")))
library(corrplot)

```

````{R}
# Calculating the correlation matrix
correlationMatrix <- cor(df_sales_num)

# Find attributes that are highly correlated
# ---
#
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.75)

# Highly correlated attributes
# ---
# 
highlyCorrelated

names(df_sales_num[,highlyCorrelated])

````



```{R}
# We can remove the variables with a higher correlation 
# and comparing the results graphically as shown below
# ---
# 
# Removing Redundant Features 
# ---
# 
Dataset2<-df_sales_num[-highlyCorrelated]

# Performing our graphical comparison
# ---
# 
par(mfrow = c(1, 2))
corrplot(correlationMatrix, order = "hclust")
corrplot(cor(Dataset2), order = "hclust")

```